### 2009-11-24 docs/marcello-dissertation/thesis/chapter7-data-persistence-component-implementation.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

This section shows the implementation of the DSP Data Persistence, which
follows the specifications of the DSP Components described on section 2.4,
along with the setup of the mongoDB, the chosen database system that handles
document-oriented instances. The implementation of the DSP Data Persistence
component was developed using the DSP Subversion repository at
http://code.google.com/p/netbeams.

\section{DSP Deployment on OSGi Framework}

    * Description of the OSGi MANIFEST.MF
    * Description of the component on the config.xml
    * Adding a matching rule on the matcher\underline{ }config.xml
    * Adding a bootstrap message the component and database configuration

    * Adding a new entry into the deployment configuration artifact config.xml;
    * Adding a new entry into the matcher configuration artifact matcher\underline{ }config.xml;
    * Adding an optional configuration message that sets up the new DSP Component;
    * Adding new Java drivers that communicates with the Database system;
    * Installing and configuring the proposed database system.

\subsection{Starting the DSP Platform}

    * command-line (Knopflerfish Framework)
    * user interface (Knopflerfish Framework)

\subsection{Stopping the the DSP Platform}

    * command-line (Knopflerfish Framework)
    * user interface (Knopflerfish Framework)

@@

nstallation of mongoDB
- Explanation about mongoDB processes: mongod, mongos, mongo
- Specification of the shards, shard key, etc

@@

\lstset{label=cmd:mongo-findone,caption=Querying the database: one item}
\begin{lstlisting}
> db.SondeDataContainer.findOne()
{"_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b" ,
"transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "45.01" , "sp_condition" : "37.6"
, "condition" : "145.8" , "resistence" : "159.77" , "salinitude" : "0.0" , "pressure" : "0.391" , "depth" : "0.46" , "ph" : "5.64" ,
"pH_mv" : "-62.1" , "odo_sat" : "89.7" , "odo_condition" : "59.34" , "turbidity" : "0.0" , "battery" : "9.4"}}
\end{lstlisting}

>>>>

This section shows the implementation of the DSP Data Persistence, which
follows the specifications of the DSP Components in the previous chapter,
along with the setup of the mongoDB, the chosen database system that models
data in a type of key-value pair model. Moreover, this chapter also describes
the experiments conducted for the persistence.

The implementation of the DSP Data Persistence component, as well as the
implementation of the experiments, were developed and managed using the DSP a
reserved development branch at the Subversion \cite{subversion} repository
located at the Google Code website. In this way, anyone can checkout the entire
source-code from the branch
\textit{https://netbeams.googlecode.com/svn/branches/marcello/persistence},
as shown in Figure \ref{fig:dsp-data-persistence-dir}. In this way, the
recommended checkout branch directory will give the remaining directory
structure as summarized:

\begin{itemize}
  \item \textbf{NEBEAMS-DIR/}: the directories under the subdirectory
  ``persistence'';
  \item \textbf{DEV-VERSION}: the directory of the current development
  version for the DSP; That is, ``\textbf{NEBEAMS-DIR}/versions/v2/''
  \item \textbf{PERSISTENCE-DIR}: the directory of the DSP component, that is,
  ``\textbf{DEV-VERSION}/apps/osgi-intro-bundles/dsp/DSPDataPersistence''.
\end{itemize}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.6]{../diagrams/dsp-data-persistence-dir}
  \caption{The DSP Data Persistence Directory Structure}
  \label{fig:dsp-data-persistence-dir}
\end{figure}

The DSP Platform defines its own run-time directory, which organizes the
artifacts used by the osgi-intro Framework, as well as the its own configuration
descriptor and matcher, as described in the previous section.

\begin{itemize}
  \item \textbf{RUNTIME-DIR}: the main directory for the execution of the DSP
Platform, with all the necessary osgi-intro-related artifacts are placed, as shown in
Figure \ref{fig:dsp-runtime-dir}.
\end{itemize}

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/dsp-runtime-dir}
  \caption{The DSP Data Persistence Directory Structure}
  \label{fig:dsp-runtime-dir}
\end{figure}

\section{DSP Platform Deployment}

As described in the previous chapter, the DSP Data Persistence Component was
implemented using the Java Programming language, on top of the osgi-intro Framework
API. This section describes the setup process of the implementation, as well as
the artifacts use during the implementation, which will be listed in the
appendix section.

The structure of the artifacts implemented for the DSP Data Persistence
Component follow the conventions of the NetBEAMS implementation, where the main
directory \textbf{PERSISTENCE-DIR}, depicted by Figure
\ref{fig:dsp-data-persistence-dir-checkedout}, holds each of the development
artifacts.

\begin{itemize}
  \item \textbf{DSPDataPersistence}: Main directory with the build.xml (Listing
  \ref{file:dsp-build.xml}) and other Eclipse-related artifacts. The other
  major directories are also listed:
  \item \textbf{META-INF}: this directory contains the descriptor file for
  the osgi-intro Framework, this this case the MANIFEST.MF;
  \item \textbf{src}: the main directory structure for the source-code
  implemented. Note that it follows the Java specification for packaging, and
  therefore, lists the package org.netbeams.dsp.persistence as the main
  package, including the layers controller, model and osgi-intro.
\end{itemize}

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/dsp-data-persistence-dir-checkedout}
  \caption{The DSP Data Persistence Directory Structure}
  \label{fig:dsp-data-persistence-dir-checkedout}
\end{figure}

\subsection{osgi-intro Deployment Process}

Since each NetBEAMS component is managed by an osgi-intro component and its
infrastructure, this section describes the basic functionality of the osgi-intro
platform \cite{osgi-intro}, a framework was conceived to support modularity in
terms resources-limited environments such mobile devices and vehicles, but it
was first widely deployed on Eclipe IDE\footnote{Integrated Development
Environment}, because it promotes the use of reusable loosely-coupled modules
using the Java Programming Language \cite{java}. The most basic layer of the
osgi-intro Framework depicted in Figure \ref{fig:layering-osgi-intro} are as follows:

\begin{itemize}
  \item \textbf{Module Layer}: manages the osgi-intro bundles deployed on the osgi-intro
  Platform, providing the necessary "wiring" of the components. In other words,
  the modules, called osgi-intro Bundles, can export and/or import packages in the
  level of a Java Class managed by the osgi-intro Framework;
  \item \textbf{Service Layer}: resposible for the interoperability between 2
  or more bundles, enabling the bundles to register services offered by
  the its specification;
  \item \textbf{Execution Layer}: executes the bundles, managing and changing
  their life-cycle through the bundle execution.
\end{itemize}

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/layering-osgi}
  \caption{The osgi-intro Framework Layers}
  \label{fig:layering-osgi-intro}
\end{figure}

NetBEAMS uses the concept of the modularization through the use of the the
Producer-Consumer paradigm described in the previous section. Since the DSP
Components are essentially osgi-intro bundles, the interoperability between them are
given in the implementation and reuse existing bundles in the osgi-intro Framework
and the DSP Platform, described by an artifact called MANIFEST.MF as seen in
listing \cite{file:osgi-intro-manifest}. In general, an osgi-intro bundle must
provide specifications that describes the module to be published into the osgi-intro
Framework. In this way, the main properties of the osgi-intro MANIFEST.MF artifact can
used by the DSP Data Persistence can be summarized as follows:

\begin{itemize}
  \item \textbf{Bundle-Activator}: The name of the instance of an osgi-intro Activator
  class, responsible to manage the bundle. The implemented class
  ``DataPersistenceActivator'' provides the activation mechanisms for the
  services provided by the DSP Data Persistence Component;
  \item \textbf{Bundle-ClassPath}: the necessary Java Jars list needed to run
  the bundle. It lists the mongoDB Java driver as one of the required
  dependencies deployed in the same package;
  \item \textbf{Import-Package}: the Java Packages needed by the DSP Data
  Persistence Component Bundle. These Java Packages must be provided by other
  DSP components, forcing the execution to be dependent on those package.
  They are provided through a ``Export-Package'' section in other osgi-intro
  bundles and, as it is shown in listing \cite{file:osgi-intro-manifest}, they
  are from different packages.
\end{itemize}

Once the osgi-intro bundle is installed into the osgi-intro Platform, it will be managed by
the osgi-intro Execution layer and change the bundle state according to its
lify-cycle, as described in the previous chapter. Therefore, the packaging of
the osgi-intro source-code is done using a JAR\footnote{Java Archival Repository}
artifact specification \cite{java-tutorial}. In this way, the DSP Data
Platform osgi-intro bundle JAR is created by the Apache ANT build script
\cite{apache-ant} in Listing \ref{file:dsp-build.xml}. The task
``dsp-data-persistence.all'' compiles all the source-code created from the
design of the previous chapter and packages everything into the artifact
``DSPDataPersistence-x.x.jar'', where x.x is the version of the bundle. Figure
\ref{fig:dsp-runtime-dir} shows the DSP Data Persistence file under the
dirctory ``\textbf{RUNTIME-DIR}/deployment/''.

\subsection{Adding DSP Data Persistence into DSP Platform }

Once the new component was developed, the DSP Data Persistence component was
deployed by editing two main descriptors, as well as the optional bootstrap
message, as described in chapter 5:

\begin{itemize}
  \item \textbf{config.xml}: the DSP Data Persistence is added to the DSP
  Platform adding an entry for the component;
  \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
  filters the messages to the DSP Data Persistence;
  \item \textbf{DSPDataPersistence\underline{ }001.xml}: depicts the bootstrap
  message for the DSP Data Persistence Component, as shown in Listing 
  \ref{file:dsp-data-pers-bootstrap.xml}.
\end{itemize}

The first configuration artifact enables NetBEAMS to initialize contain the
descriptor of the component, as shown in Listing \ref{file:dsp-config.xml}.
The component is added with the highest priority, since it contains
dependencies to other DSP Components. 
\subsection{Starting the DSP Platform}

Once the DSP Platform automatically installs all the components described by
the configuration descriptor, the container verifies the dependencies and
starts each of the DSP components of Figure \ref{fig:knopflerfish-execution},
the osgi-intro Framework container Knopflerfish \cite{knopflerfish}.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.65]{../diagrams/knopflerfish-execution}
  \caption{The execution of the Knopflerfish Container}
  \label{fig:knopflerfish-execution}
\end{figure}

@@

The artifacts of the mongoDB are located in the thirdparty directory of the
NetBEAMS resources. However, the build script from the DSP Data Persistence
component can be executed to produce the persistence directory
structure for NetBEAMS, as described in figure
\ref{fig:dsp-persistence-system-dir}.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/dsp-persistence-system-dir}
  \caption{The osgi-intro Framework Layers}
  \label{fig:dsp-persistence-system-dir}
\end{figure}

Installation of mongoDB
- Explanation about mongoDB processes: mongod, mongos, mongo
- Specification of the shards, shard key, etc

@@

\lstset{label=cmd:mongo-findone,caption=Querying the database: one item}
\begin{lstlisting}
> db.SondeDataContainer.findOne()
{
    "_id" : ObjectId("5d6f40078ec8074bba980900"),
    "message_id" : "020d82e1-18b2-4fe2-800c-a0f68e22ea86",
    "sensor" : {
        "ip_address" : "192.168.0.91",
        "location" : {
            "latitude" : 37.89155,
            "longitude" : -122.4464
        }
    },
    "time" : {
        "valid" : "Wed Nov 04 2009 02:30:12 GMT-0800 (PST)",
        "transaction" : "Sat Nov 21 2009 03:01:33 GMT-0800 (PST)"
    },
    "observation" : {
        "WaterTemperature" : 88.58,
        "SpecificConductivity" : 180.5,
        "Conductivity" : 167.3,
        "Resistivity" : 491.89,
        "Salinity" : 0.06,
        "Pressure" : 1.109,
        "Depth" : 2.642,
        "pH" : 7.11,
        "pHmV" : -42.9,
        "Turbidity" : 0.2,
        "ODOSaturation" : 72.4,
        "ODO" : 10.85,
        "Battery" : 2.7
    }
}
\end{lstlisting}
